---
title: "Unsupervised Learning"
date: "3/6/2020"
output: html_document
---
## 1. Principal Component
We use the 'USArrests' data

```{r}

dimnames(USArrests)
apply(USArrests, 2, mean) # second parameter 1 - refers ROWS, 2 - refers COLUMNS
apply(USArrests, 2, var)


```

We see that 'Assualt' has much larger variance than the other variables. It would dominate the prinicpal components. So we need to standeradize when perform PCA. 

```{r}
pca.out = prcomp(USArrests, scale=TRUE) # perform a principal component analysis. 
pca.out

```
```{r}
names(pca.out)
pca.out
biplot(pca.out, scale = 0, cex=0.7)
```

## 2.K-Means Clustering
  K-Means works in any dimension but we have chosen 2 dimension and made some data.
  
```{r}
set.seed(101)
x = matrix(rnorm(100*2), 100, 2)
xmean = matrix(rnorm(8, sd=4), 4, 2)
which = sample(1:4, 100, replace = TRUE)
x = x + xmean[which, ]
#plot and assign the initial cluster Ids
plot(x, col=which, pch=19)
```

  
We know the 'true' cluster Ids, but will not tell that to K-Means Alorithm.

```{r}
km.out = kmeans(x,4, nstart = 15) #nstart ?
km.out$cluster
plot(x, col=km.out$cluster, cex=2, pch=1, lwd=2)
points(x, col=which, pch=19)
points(x, col=c(4,3,2,1)[which],pch=19 )
```

(3) Hierarcical Clustering

we will use the same data & use hierarchical clustering. 

```{r}
hc.complete = hclust(dist(x), method="complete")
plot(hc.complete, cex=2)
hc.single = hclust(dist(x), method="single")
plot(hc.single)
hc.average = hclust(dist(x), method="average")
plot(hc.average)

```


 
Lets compare this withthe actual clusters in the data. We will use the function 'cuttree' to cut the tree at level 4. This funciton produce a vector of numbers from 1 to 4, saying which branch each observation is on. sometiems you will see pretty plots where the leaves of teh dendrogram are colored. 

We can use 'table' to see how to well they match. 

```{r}
hc.cut = cutree(hc.complete, 4)
table(hc.cut, which)
table(hc.cut, km.out$cluster)
```


Or, we can use our group membership as labels for te leaves of te dendrogram. 

```{r}
plot(hc.complete, labels = which)
```


