---
title: "Ch7_NonLinearModel"
author: "Hencil Peter"
date: "2/15/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Non-Linear Model
================
Polynomials:
  - First we will use polynomials and focus a single predictor "age"
  

```{r}
require(ISLR)
attach(Wage)

fit = lm(wage~poly(age,4), data = Wage)
summary(fit)
#above polynomial 4 is not very significant 
```
The "poly" function generates a basis of *orthogonal polynomials*
Lets make a plot of the fitted function, along with the standard error of the fit.

```{r fig.width=7, fig.height=6}
agelims = range(age)
age.grid = seq(from=agelims[1], to=agelims[2])
preds = predict(fit, newdata = list(age=age.grid),se=TRUE)
se.bands = cbind(preds$fit+2*preds$se, preds$fit-2*preds$se)
plot(age, wage, col="darkgrey")
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, col="blue", lty=2)
```
There are other direct ways of doing this in R.
e.g 
```{r}
fita = lm(wage~age+I(age^2)+I(age^3)+I(age^4), data=Wage)
summary(fita)
```
By using orthogonal polynomials in this simple way, it turns out that we can separately test for each coefficients. 
So if we look at the summary, linear, quadratic & cubic terms are significant, but not the quartic.

Further, this only works with linear regression with single predictor. we will use "anova" in the next demonstration. 

```{r}
fita = lm(wage~education, data = Wage)
fitb = lm(wage~education+age, data = Wage)
fitc = lm(wage~education+poly(age,2), data = Wage)
fitd = lm(wage~education+poly(age,3), data = Wage)
anova(fita, fitb, fitc, fitd)
```

Polynomial Logistic Regression
==============================

Now we fit a logistic regression modelto a binary response variable, constructed from "wage". we code teh big earns (">250K") as 1 else 0.

```{r}
fit = glm(I(wage>250)~poly(age,3), data = Wage, family = binomial)
summary(fit)
preds = predict(fit, list(age=age.grid), se=T)
se.bands = preds$fit + cbind(fit=0, lower=2*preds$se, upper=2*preds$se)
se.bands[1.5,]

```
We have done the computations on the logit scale. To transform we need to apply inverse logit mapping. 

we can do this simultaniously for all three columns of "se.bands".

```{r}
prob.bands = exp(se.bands)/(1+exp(se.bands))
matplot(age.grid, prob.bands, col="blue", lwd=c(2,1,1), lty=c(1,2,2), type="l", ylim=c(0,1))
points(age, I(wage>250)/10, pch="|", cex=.5)
```

Splines
=======
- is a cubic polynomial (function bs used). idea is more local then polynimial.
- they construct usign continious knots 

Cubic Splines
============
```{r}
require(splines)
fit = lm(wage~bs(age, knots=c(25,40,60)), data=Wage)
plot(age, wage, col="darkgrey")
lines(age.grid, predict(fit, list(age=age.grid)), col="darkgreen", lwd=2)
abline(v=c(25,40,60), lty=2, col="darkgreen")  
summary(fit)
```

Smoothing Splines
================
-Does not require knot selection but it does have smoothing parameters 
- Which can be specified via degrees of freedom "df"

```{r}
fit = smooth.spline(age,wage, df=16)
#Below line expect plot to be called before. ERROR. Hencil to fix. 
lines(fit, col="red", lwd=2)
```
Use LOO cross validation to select smoothing prarameter automatically. 

```{r}
fit = smooth.spline(age, wage, cv=TRUE)
#below line error 
lines(fit, col="purple", lwd=2)
fit 
```

Generalized Additive Models
===========================

So far we have focused on fitting models with mostly  single linear non-linear regression. The "gam" package makes it easier to work with multiple non-linear terms. In addition, it knows how to plot these functions and their standard error. 

```{r fig.width=10, fig.height=5}
#install.packages("gam")
require(gam)
gam1=gam(wage~s(age, df = 4)+s(year, df=4)+education, data=Wage)
par(mfrow=c(1,3))
plot(gam1, se=T)
gam2=gam(I(wage>250)~s(age, df = 4)+s(year, df=4)+education, data=Wage, family = binomial)
plot(gam2)
```
Non-Linear term for year. 
```{r fig.width=10, fig.height=5}
gam2a = gam(I(wage>250)~s(age, df=4)+year+education, data=Wage, family=binomial)
anova(gam2a, gam2, test="Chisq")
```

gam package knows how to plot the functions nicely, even for model fit by 'lm' and 'glm'

e.g 
```{r fig.width=10, fig.height=5}
par(mfrow=c(1,3))
lm1=lm(wage~ns(age, df=4)+ns(year,df=4)+education, data=Wage)
plot.Gam(lm1, se=T)

```


